---
title: "Task 24"
author: "Logan Herrmeyer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
pacman::p_load(
  tidyverse, dygraphs, DT, tidyquant, timetk, lubridate)
kr <- tidyquant::tq_get("KR") %>%
  filter(date >= today() - years(5))
```

## Kroger Stock Price

```{r}
dygraph(xts(kr$adjusted,kr$date),
        main="Kroger stock price")
```

## Kroger investment returns

```{r}
dygraph(xts(kr$adjusted,kr$date), 
        main="Value of $10,000 invested in Kroger in 2016") %>%
    dyRebase(10000) %>%
    dyAnnotation("2017-06-14","A") %>%
    dyAnnotation("2017-09-12","B")
```
A: [Stock price drops](https://www.fool.com/investing/2017/07/05/why-the-kroger-co-stock-plunged-217-in-june.aspx) because earnings were only average even with increased revenue, and Amazon acquired Whole Foods.

B: [Stock price drops](https://www.cnbc.com/2018/09/13/kroger-shares-crash-as-sales-dissapoint.html) because sales were less than expected.

## Summary

The first graph shows the stock price of Kroger over the last five years, and the second shows the value of $10,000 invested in Kroger over the past five years, with points of significant price movement. I picked those points because those are the points where I could find information on the cause of the price movement.