---
title: "Task 6"
author: "Logan Herrmeyer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Task 6: Fix This Chart

Here is the original chart. I have commented each line of code in the chart to show what it does.
```{r, message = FALSE}
library(readr) # Import CSV reading library
library(ggplot2) # Import graphing library

rcw <- read_csv("https://byuistats.github.io/M335/data/rcw.csv", # Read the CSV from the internet
                col_types = cols(Semester_Date = col_date(format = "%m/%d/%y"), # Specify the date format for the date column
                                 Semester = col_factor(levels = c("Winter", "Spring", "Fall")))) # Give the semester column 3 options, Winter, Spring, and Fall
head(rcw) # Get the first few rows of the dataset and print them
 
ggplot(data = rcw, # Plot with the RCW dataset
       mapping = aes(x = Semester_Date, # Set the X axis to the semester
                     y = Count, # Set y axis to number of people attending
                     color = Department)) + # Set color to be based on department.
  geom_line() + # Make the graph a line graph
  geom_point() # Make the graph also a point graph
```

In this graph, there are definite trends. The ME department fairly consistently has the highest attendance, but it did suffer a decline in 2017 followed by a recovery. Mid-2017 seemed to be a low point for all departments. The Math and DCM departments have consistently had the lowest attendance.

I have made my own graph to represent the data, shown here:
```{r}
ggplot(data = rcw, mapping = aes(fill=Department,x=paste(Semester,Year),y=Count)) + geom_bar(stat="identity")
```

I used this style of chart because it helps to show the relative popularity of one group compared to another for each semester. I used the new function geom_bar() from the readings to make this graph.